Interfering with Noninterference \iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF and...?)}}\fi

### 2020-02-02 main.tex
Comments

Concretely, we have in mind a conventional machine architecture with a
RISC-V processor, perhaps extended with some kind of hardware-enforced
security policy, but we won't need any RISC-V-specific details till we get
to \cref{impl}.

@@

\leo{We might need to change the intuition behind some of these things.}
A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow
\{\HIGH,\LOW\} \times \{\HI, \LO\}$ is a map from machine components
to confidentiality and integrity levels. $\LOW$ marks locations whose
contents are observable from the outside (by the ``attacker,'' e.g.,
by DMA) and $\HIGH$ marks internal ``hidden'' state that cannot be
observed directly (e.g., processor registers and memory addresses for
temporary variables will usually be hidden). Similarly, $\LO$
locations correspond to low-integrity components (e.g. locations that
can be written to by the network), while $\HI$ locations should be
protected from such flows.

@@

A {\em trace} of length $n$, where $n$ can be finite or infinite, is a
length-$n$ sequence $\seq{\mach}_n = \mach_0,\, \mach_1,\ldots$ of
machine states in which adjacent machine states are related by
$\stepstounderfull$---i.e., in which $\mach_{i-1} \stepstounderfull
\mach_i$ for each $i > 0$ in the sequence.

A {\em contoured hyperproperty} is a set of triples: a contour along
with a pair of traces.

@@

Finally, we can parameterize the traces by some initial
application-specific state to obtain a {\em contoured lockstep
  $X$-hyperproperty}, where every trace comes with more information,
for example call sites in stack safety.

@@

  For each triple $(\contour, \overline{M}_n, \overline{M'}_{n'})$, if
  the initial elements of the traces are variants, then the final
  states of the traces are also variants.

@@

\leo{To warm up, a super simple separation policy.}
\begin{itemize}
\item Any instruction can serve as a final instruction.\leo{What about
  instructions the are halted by the monitor?}
\item For forward compatibility we could treat the first unmatched return as
  a halt instruction.
\item The instruction memory is marked $\LOW \times \HI$. That means
  it is observable and can't change.
\item The memory space is split in two. One part is marked $\LOW
  \times \LO$: this is the public part (or the one that can be
  accessed by the current process e.g.). The other part is marked
  $\HIGH \times \HI$: it should be completely inaccessible to the
  current process.
\item Separation is noninterference:
  \leo{Exactly the definition above}
\item Depending on what we do with final instructions, the property
  could be parameterized by the ``call/return sites''.
\end{itemize}
\leo{(The $+$ part in the title is a hint that one thing that gets added is that
the PCs of corresponding traces always stay in lock step because of integrity?)}

@@

\leo{Build up to the interesting thing:}
\begin{itemize}
\item The first unmatched return serves as a halt.
\item The instruction memory is again marked $\LOW \times \HI$. That
  means it is observable (both traces have the same instructions!) and
  can't change.
\item At the beginning of time, the entire data memory is marked $\LOW
  \times \LO$. That makes for a very unintersting property as all
  variants are necessarily identical.
\item BUT, we close the hyperproperty as follows:
  \begin{itemize}
  \item We need the state description stuff to enforce WBCF and be
    able to talk about ``first unmatched returns''. That is also a
    trace property.
  \item After every call, we keep the part of the (left-hand, wlog
    due to symmetry) trace until its first unmatched return.
  \item We construct a new contour at that point: after every call,
    the local state of the caller becomes inaccessible (marked
    $\HIGH \times \HI$) until a subsequent return.
  \item For every variant of the new initial state with respect to the
    new contour, the contour plus pair of traces is in the
    hyperproperty.
  \end{itemize}  
\end{itemize}

>>>>

Concretely, we have in mind a conventional machine architecture with a
standard (RISC-V) processor extended with some kind of (static or dynamic,
software- or hardware-enforced) security monitor, but we won't need to make
any specific choices about details until we get to \cref{impl}.

@@

\leo{We might need to change the intuition behind some of these things.}
A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow
\{\HIGH,\LOW\} \times \{\HI, \LO\}$\bcp{We need more mnemonic names for
  these (that help us keep straight when we're talking about confidentiality
  and when integrity), both in the tex and in the way they are displayed.}
is a map from machine components 
to confidentiality and integrity levels. $\LOW$ marks locations whose
contents are observable from the outside (by the ``attacker,'' e.g.,
by DMA) and $\HIGH$ marks internal ``hidden'' state that cannot be
observed directly (e.g., processor registers and memory addresses for
temporary variables will usually be hidden). Similarly, $\LO$
locations correspond to low-integrity components (e.g. locations that
can be written to \bcp{directly or indirectly} by the network), while $\HI$ locations should be
protected from such flows.

\bcp{A random thought about integrity: One thing that may be helping confuse
  us about integrity is that we've gone back and forth between ``eager'' and
  ``lazy'' versions of various micropolicies, where the ``lazy'' ones do
  allow high-integrity data to be overwritten, but only in a way that it
  cannot be read again later (e.g., the lazy heap policy guarantees that, if
  inaccessible memory is overwritten, then its color will change).  The lazy
point of view is much more subtle as an integrity property.  So, to keep our
heads straight here, we should be thinking only about eager properties in
the first instance.}

@@

A {\em trace} of length $n$, where $n$ can be finite or infinite\bcp{awkward}, is a
length-$n$ sequence $\seq{\mach}_n = \mach_0,\, \mach_1,\ldots$ of
machine states in which adjacent machine states are related by
$\stepstounderfull$---i.e., in which $\mach_{i-1} \stepstounderfull
\mach_i$ for each $i > 0$ in the sequence.

A {\em contoured hyperproperty} is a set of triples, each consisting of a
contour along with a pair of traces.

@@

Finally, we can parameterize the traces by some initial application-specific
state to obtain a {\em contoured lockstep $X$-hyperproperty}, where every
trace comes with more information, for example call sites in stack safety.
\bcp{If we use this later, we need to say it more formally to be clear about
  what we mean---e.g., I had thought both the traces and the contour would
  be functions of $X$!}

@@

  For each triple $(\contour, \overline{M}_n, \overline{M'}_{n'})$, if
  the initial elements of the traces are variants, then the final
  states of the traces are also variants.  \bcp{What does ``final''mean for
    infinite traces?  What about pairs where one is finite and the other
    infinite?}

@@

\leo{To warm up, a super simple separation policy.}
\begin{itemize}
\item Any instruction can serve as a final instruction.  \bcp{My picture
  here was that, for this property, the $X$ parameter would be, say, the
  address where the machine is considered to halt. (I guess the picture is
  that we attach probes to the PC and to the rest of memory, and we read out
  the memory when the PC reaches this address.  But when I write it down
  that way it seems pretty awkward.  Maybe we should just assume that the
  machine has a HALT instruction or some other way of signaling that it is
  done---i.e., abstractly, simply take the step function as a partial
  function.  We will need to put a footnote somewhere acknowledging that
  this may be a bit of an abstraction from real modern processors, but the
  simple separation policy is an abstraction anyway!)}

\leo{What about
  instructions the are halted by the monitor?}  \bcp{Not sure what the
  question is here, but note that an instruction halted by the monitor will
  not appear in the trace.}
\item For forward compatibility we could treat the first unmatched return as
  a halt instruction.
\item The instruction memory is marked $\LOW \times \HI$. That means
  it is observable and can't change.  \bcp{What does this assumption mean,
    precisely?  (I.e., what is ``the instruction memory''?)}
\item The memory space is split in two. One part is marked $\LOW
  \times \LO$: this is the public part (or the one that can be
  accessed by the current process e.g.). The other part is marked
  $\HIGH \times \HI$: it should be completely inaccessible to the
  current process.
\item Separation is noninterference:
  \leo{Exactly the definition above}
\item Depending on what we do with final instructions, the property
  could be parameterized by the ``call/return sites''.
\end{itemize}
\leo{(The $+$ part in the title is a hint that one thing that gets added is that
the PCs of corresponding traces always stay in lock step because of integrity?)}
\bcp{The section titles are not working so well, now that we're being a bit
  less abstract about our definitions.}

@@

\leo{Build up to the interesting thing:}
\begin{itemize}
\item The first unmatched return serves as a halt.  \bcp{What does that mean
formally?}
\item The instruction memory is again marked $\LOW \times \HI$. That
  means it is observable (both traces have the same instructions!) and
  can't change.
\item At the beginning of time\bcp{Not clear what ``beginning of time''
  means when we're talking about hyperproperties---the intuition of time is
  relevant to the traces within the hyperproperty, but not to the
  hyperproperty itself.  Simpler, I think, to view the hyperproperty as a
  starting set of traces (where the whole machine state [registers and PC too,
  right?] is marked $\LOW$), closed under an operator that takes a
  contour+trace+trace and snips out all the subtraces between a call and its
return.  (But: eek---see below.)}, the entire data memory is marked $\LOW
  \times \LO$. That makes for a very unintersting property as all
  variants are necessarily identical.
\item BUT, we close the hyperproperty as follows:
  \begin{itemize}
  \item We need the state description stuff to enforce WBCF and be
    able to talk about ``first unmatched returns''. That is also a
    trace property.
  \item After every call, we keep the part of the (left-hand, wlog
    due to symmetry) trace until its first unmatched return.
  \item We construct a new contour at that point: after every call,
    the local state of the caller becomes inaccessible (marked
    $\HIGH \times \HI$) until a subsequent return.
  \item For every variant of the new initial state with respect to the
    new contour, the contour plus pair of traces is in the
    hyperproperty.
  \end{itemize}
\end{itemize}

\bcp{Eeek.  Just realized why we absolutely need the ``lockstep'' part!
  The closure operator that snips out a ``subtrace'' between a call and the
  corresponding return has to do the snipping on {\em both} traces at once,
  and this only makes sense if they are executing in lockstep.}
